---
title: "COVID-19 Virus Trends"
author: "RF"
date: "`r Sys.Date()`"
output: html_document
---
## Project description
Basic analysis of COVID-19 trends on a country level, as a means of learning R Markdown.

### Data exploration
```{r}
library(readr)
covid_df <- read.csv('covid19.csv')
dim(covid_df)
vector_cols <- colnames(covid_df)
vector_cols
head(covid_df)
library(tibble)
glimpse(covid_df)
```
### Data cleaning
We only want data that represents a whole country. 
```{r}
library(dplyr)
covid_df_all_states <- covid_df %>% 
  filter(Province_State == "All States") %>%
  select(-Province_State)
```
We decided to work with daily data. So we need to extract columns related to daily measures.
```{r}
covid_df_all_states_daily <- covid_df_all_states %>%
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)
```
### Summarize the data
Extract the top ten cases country data.
```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>%
  group_by(Country_Region) %>%
  summarise(
    tested = sum(daily_tested),
    positive = sum(daily_positive),
    active = sum(active),
    hospitalized = sum(hospitalizedCurr)
    ) %>%
  arrange(-tested)
covid_df_all_states_daily_sum
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
```
### Positive cases against number of tests
As a means of practicing vectors.
```{r}
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
positive_cases / tested_cases
positive_tested_top_3 <- c("United Kingdom", "United States", "Turkey")
```
Save information about these top 3 countries, as a means of learning vectors and matrices.
```{r}
united_kingdom <- c(0.11,1473672,166909,0,0)
united_states <- c(0.10,17282363,1877179,0,0)
turkey <- c(0.08,2031192,163941,2980960,0)
top_3 <- rbind(united_kingdom, united_states, turkey)
colnames(top_3) <- c("Ratio", "tested", "positive", "active", "hospitalized")
top_3
```
Pull all the answers and datasets together, as a means of learning lists.
```{r}
question <- "Which countries have had the highest number of positive cases against the number of tests?"
answer <- c("Positive tested cases" = positive_tested_top_3)
dataframes <- list(covid_df, covid_df_all_states, covid_df_all_states_daily, covid_df_all_states_daily_sum, covid_top_10)
matrices <- list(top_3)
vectors <- list(vector_cols, countries)
data_structure_list <- list(dataframes=dataframes, matrices=matrices, vectors=vectors)
covid_analysis_list <- list(question, answer, data_structure_list)
covid_analysis_list[[2]]
```
