---
title: "DQ Book Sales"
author: "RF"
date: "`r Sys.Date()`"
output: html_document
---

### Data exploration
Download the necessary libraries and dataset.
```{r}
library(tidyverse)
sales <- read_csv("sales2019.csv")
```
How big is the dataset?
```{r}
dim(sales)
```
What are the column names and what do they represent?
```{r}
colnames(sales)
```
date: date 
user_submitted review: review submitted by user
title: title of book
total_purchased: total number of copies purchased by this customer
customer_type: business or individual

What are the column types?
```{r}
for (c in colnames(sales)) {
  print(c)
  print(typeof(sales[[c]]))
}
```
Do any of the columns have missing data?
```{r}
for (c in colnames(sales)) {
  print(c)
  print(sum(is.na(sales[[c]])))
}
```
### Numeric data cleaning
Remove rows with missing user_submitted_review data.
```{r}
clean_sales <- sales %>%
  filter(!is.na(user_submitted_review))
dim(clean_sales)
```
Calculate average number of books purchased on an order.
```{r}
avg_num_purchased <- mean(clean_sales$total_purchased, na.rm = TRUE)
avg_num_purchased
```
Fill missing values in total_purchased with the average value calculated above.
```{r}
clean_sales <- clean_sales %>%
  mutate(
    clean_total_purchased = if_else(is.na(total_purchased), avg_num_purchased, total_purchased)
  )
```
### String data cleaning
Words or phrases that indicate a positive review: "Awesome", "OK", "Never", "a lot".
Create a function that indicates whether or not a review contains one of these words or phrases.
```{r}
positive_review <- function(review) {
  case_when(
    str_detect(review, "Awesome") ~ TRUE,
    str_detect(review, "OK") ~ TRUE,
    str_detect(review, "Never") ~ TRUE,
    str_detect(review, "a lot") ~ TRUE,
    TRUE ~ FALSE
  )
}
```
Create a new column that indicates whether or not the review is positive.
```{r}
clean_sales <- clean_sales %>%
  mutate(
    is_positive = positive_review(user_submitted_review)
  )
```
### Data Analysis
Was the new program effective in increasing book sales? The program started on July 1, 2019.

Convert date column using lubridate.
```{r}
clean_sales$date <- lubridate::ymd(clean_sales$date)
```
Group sales that happen before and after the program started.
```{r}
program <- lubridate::ymd("2019-07-01")
clean_sales <- clean_sales %>%
  mutate(
    before_after = case_when(
      date < program ~ "before",
      TRUE ~ "after"
    )
  )
```
Create a summary table that compares the number of books purchased before and after the program started.
```{r}
clean_sales %>%
  group_by(before_after) %>%
  summarise(
    books_purchsed = sum(clean_total_purchased)
  )
```
The program did not increase overall sales. Let's break down the analysis by customer_type to see if there was a change there.
```{r}
clean_sales %>%
  group_by(before_after, customer_type) %>%
  summarise(
    books_purchsed = sum(clean_total_purchased)
  )
```
It looks like businesses purchased more but individuals purchased less.

How did the program impact review scores? Did review scores improve?
```{r}
clean_sales %>%
  group_by(before_after) %>%
  summarise(
    num_positive_reviews = sum(is_positive)
  )
```
The program didn't seem to change the number of positive reviews.