---
title: "DQ SQL Business Questions"
author: "RF"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setup
Import the `RSQLite` and `DBI` libraries. Then, create helper functions `run_query()` which returns the result of an SQL query and `show_tables()` that calls `run_query()` and returns a list of tables and views.

```{r}
library(RSQLite)
library(DBI)

run_query <- function(query, db) {
  conn <- DBI::dbConnect(SQLite(), db)
  result <- DBI::dbGetQuery(conn, query)
  DBI::dbDisconnect(conn)
  return(result)
}

show_tables <- function(db) {
  query <- 
  "SELECT
    name,
    type
  FROM sqlite_master
  WHERE type IN ('table','view');"
  tables <- run_query(query, db)
  return(tables)
}

db <- 'chinook.db'
show_tables(db)
```
### Which Generas Sell the Most in the US?
Write a query that for each genre, calculates the number of tracks sold in the US in absolute numbers and percentages.
```{r}
query <-
  "WITH usa_tracks_sold AS
   (
    SELECT il.* FROM invoice_line il
    INNER JOIN invoice i on il.invoice_id = i.invoice_id
    INNER JOIN customer c on i.customer_id = c.customer_id
    WHERE c.country = 'USA'
   )
SELECT
    g.name genre,
    count(uts.invoice_line_id) tracks_sold,
    cast(count(uts.invoice_line_id) AS FLOAT) / (
        SELECT COUNT(*) from usa_tracks_sold
    ) percentage_sold
FROM usa_tracks_sold uts
INNER JOIN track t on t.track_id = uts.track_id
INNER JOIN genre g on g.genre_id = t.genre_id
GROUP BY 1
ORDER BY 2 DESC
LIMIT 10;
      "

result <- run_query(query, db)
```

Create a plot to show the resulting table.

```{r}
library(tidyverse)

result %>%
  ggplot(aes(x = genre, y = tracks_sold, fill = genre)) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("Number of Tracks Sold")

result %>%
  ggplot(aes(x = genre, y = percentage_sold, fill = genre)) +
  geom_col() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("Percentage of Tracks Sold")
```

Rock is by far the most purchased genre in the US, by both absolute number and percentage.

### Employee Performance

Determine performance of sales support agents by analyzing purchases of customers belonging to each agent.

First, write query to pull total dollar amount of sales assigned to each sales support agent.

total dollar amount of sales assigned to each sales support agent within the company

```{r}
performance_query <- "
WITH customer_invoice AS
  (
  SELECT
    c.customer_id,
    c.support_rep_id,
    SUM(i.total) total_by_employee
  FROM customer c
  INNER JOIN invoice i ON i.customer_id = c.customer_id
  GROUP BY c.support_rep_id
  )
  
SELECT 
  e.*,
  ci.total_by_employee
FROM employee e
INNER JOIN customer_invoice ci ON ci.support_rep_id = e.employee_id
ORDER BY ci.total_by_employee DESC
"

performance_result <- run_query(performance_query, db)
performance_result
```

Plot the results of the query.
```{r}
performance_result %>%
  ggplot(aes(x=last_name, y=total_by_employee, fill=last_name)) +
  geom_col() +
  ggtitle("Employee Total Sales") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5))
```

All employees that made the chart are based in Calgary, Canada. Jane Peacock has the most sales but also the earliest hire date.

### Sales Data by Country

Write a query that collates data on purchases from different countries. For each country, include:
* total number of customers
* total value of sales
* average value of sales per customer
* average order value

Where a country has only one customer, collect them into an "Other" group.

The results should be sorted by the total sales from highest to lowest, with the "Other" group at the very bottom.
```{r}
country_query <- "
WITH customer_sales AS
  (
  SELECT
    SUM(i.total) total_per_customer,
    i.customer_id,
    c.country
  FROM invoice i
  INNER JOIN customer c ON c.customer_id = i.customer_id
  GROUP BY i.customer_id
  ),

  country_sales AS
    (
    SELECT
      CASE
        WHEN (
              SELECT count(*)
              FROM customer
              WHERE country = cs.country
              ) = 1 THEN 'Other'
          ELSE cs.country
        END AS country,
      COUNT(DISTINCT i.customer_id) customers_per_county,
      SUM(i.total) total_per_country,
      AVG(cs.total_per_customer) country_avg_per_customer,
      AVG(i.total) country_avg_order_value
    FROM invoice i
    INNER JOIN customer_sales cs ON cs.customer_id = i.customer_id
    GROUP BY 1
    ORDER BY 3 DESC
    ),
  
  country_sort AS
    (
    SELECT
      country,
      CASE
        WHEN country = 'Other' THEN 1
        ELSE 0
      END AS sort
    FROM country_sales
    )
    
SELECT cou.* FROM country_sales cou
LEFT JOIN country_sort sor ON sor.country = cou.country
ORDER BY sor.sort ASC

"

country_result <- run_query(country_query, db)
country_result
```
Brazil has only 5 customers compared to 13 in the US, but each customer spends more than in the US. This would be a good place to focus on growth.

### Album Purchases

We want to figure out what percentage of purchases are individual tracks vs whole albums. 

First, write a query that categorizes each invoice as either an album purchase or not and generates summary statistics on number and percentage.
```{r}
album_query <- "
WITH whole_invoice AS
  (
  SELECT 
    i.*,
    il.*,
    t.*
  FROM invoice i
  LEFT JOIN invoice_line il ON il.invoice_id = i.invoice_id
  LEFT JOIN track t ON t.track_id = il.track_id
  ),

is_album AS 
  (
  SELECT
    CASE
      WHEN (
        SELECT track_id 
        FROM track
        WHERE album_id = wi.album_id
        EXCEPT
        SELECT track_id
        FROM invoice_line
        WHERE invoice_line_id = wi.invoice_line_id
      ) IS NULL
      AND (
        SELECT track_id
        FROM invoice_line
        WHERE invoice_line_id = wi.invoice_line_id
        EXCEPT
        SELECT track_id
        FROM track
        WHERE album_id = wi.album_id
      ) IS NULL
      THEN 1
      ELSE 0
    END AS is_album,
    wi.invoice_id
  FROM whole_invoice wi
  )

SELECT
  SUM(ia.is_album) number_albums_perchased,
  COUNT(DISTINCT i.invoice_id) - SUM(ia.is_album) number_nonalbums_purchased,
  (SUM(ia.is_album) / CAST(COUNT(DISTINCT i.invoice_id) AS FLOAT)) *100 percent_albums_purchased,
  ((COUNT(DISTINCT i.invoice_id) - SUM(ia.is_album)) / CAST(COUNT(DISTINCT i.invoice_id) AS FLOAT)) * 100 percent_nonalbums_purchased
FROM invoice i
LEFT JOIN is_album ia ON ia.invoice_id = i.invoice_id
"
album_result <- run_query(album_query, db)
album_result
```
Almost 16% of sales are album purchases. Therefore, it does not make sense to remove albums from the platform.